---
title: "Cars_Data_Visualisation_R"
author: "Dara Abb"
date: "2025-04-10"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(tidyverse)
library(ggplot2)
```

Import datasets


```{r}
cars <- read.csv("cars.csv")

retail_clean <- read.csv("retail_clean.csv")
```

Scatterplots:are used to examine the relationship between two continuous variables

1st Scatterplot: examines the relation between horsepower an city miler per gallon 

```{r}
cars %>% ggplot(aes(x= horsepower, y=city_miles_per_galloon, color=Price))+geom_point()
```
```{r}

```


2nd Scatterplot: Weight vs. Price with Size and Color Modifications


```{r}
cars %>% ggplot(aes(x=weight, y= Price, color=sports_car, size=horsepower))+geom_point()

```

3rd Scatterplot: Scatterplot with Different Shape

```{r}
cars %>% ggplot(aes(x=weight, y= Price, shape=sports_car))+geom_point()+theme_classic()

```

4th Scatterplot: Price vs. Horsepower with Cylinders Factor

```{r}
cars %>% ggplot(aes(x=Price, y= horsepower, color=factor(cylenders)))+geom_point()
```

Bar plots help to visualize categorical data by representing the frequency of each category.

Top 9 Countries with Highest Appearances

```{r}
top9_international <- retail_clean %>% group_by(Country) %>% summarise(appearnece = n()) %>% arrange(desc(appearnece))%>% 
  slice(2:10) 

top9 <- as.vector(top9_international[["Country"]])

retail_clean %>% filter(Country %in% top9) %>% ggplot(aes(x= Country))+geom_bar(fill= "darkred")
```

Weekday Distribution Plot

```{r}
retail_clean<- retail_clean %>% separate(col = InvoiceDate,
                                         into = c("date","time"),
                                         sep = " ")

head(retail_clean)

#####weekday analysis
retail_clean$date <- as.Date(retail_clean$date)

retail_clean$week <- weekdays(retail_clean$date)

retail_clean %>% ggplot(aes(x=week))+geom_bar()
```

Bar Plot for Top 9 Countries Reordered

```{r}
top9_international %>% 
  ggplot(aes(x=reorder(Country, appearnece), y=appearnece))+geom_col()+coord_flip()+xlab("Top 9 International")

```

Density Plot: Horsepower by Cylinders

```{r}
cars %>% ggplot(aes(x=horsepower, fill = factor(cylenders)))+ geom_density(alpha=0.3)

```

Box plots: Box plots are used to visualize the distribution of a continuous variable and identify outliers.

```{r}
##Box Plot: Horsepower by Cylinders
table(cars$cylenders)
cars %>% filter(cylenders %in% c(4,6,8)) %>% ggplot(aes(x=factor(cylenders), y=horsepower))+geom_boxplot()

```

